---
title: "Access Arm Treasure Data from R with RPresto and RTD"
output:
  html_document:
    df_print: paged
---

This is an example notebook to connect Arm Treasure Data (TD) from R. If you want to know how to access from R with RJDBC or RPresto, you can see [How to connect Arm Treasure Data from R](https://rpubs.com/chezou/TD-from-R).

While RPresto enables you to query TD easily, it can't handle some TD execution. In this notebook, I will explain how to query with RPresto and how to execute basic TD operation.

Before installing R packages, you have to install [embulk](https://www.embulk.org/) and [embulk-output-td](https://github.com/treasure-data/embulk-output-td) plugin. Ensure PATH for `embulk` command to be set appropriately. You also have to set your "TD_API_KEY" is set as an environement variable.

Let's install RPresto and RTD packages.

```{r eval=FALSE}
install.packages(c("RPresto", "devtools", "dplyr"))
devtools::install_github("chezou/RTD")
```

We will use nycflights13 package for analysis. Copy flights data to TD.

```{r eval=FALSE}
install.packages("nycflights13")
```

```{r}
library(RTD)

client <- Td(apikey = Sys.getenv("TD_API_KEY"), endpoint="api.treasuredata.com")
td_upload(client, "test", "flights", nycflights13::flights, overwrite=T)
```


```{r}
library(RPresto)
library(dplyr)

db <- src_presto(
  host="https://api-presto.treasuredata.com",
  port=443,
  user=Sys.getenv("TD_API_KEY"),
  schema='test',
  catalog='td-presto'  
)

flights_tbl <- tbl(db, 'flights')
```

Execute simple dplyr operation.

```{r}
# filter by departure delay
flights_tbl %>% filter(dep_delay == 2)
```

["Introduction to dplyr"](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) introduces advanced dplyr examples. Let's try grouped operations example from the tutorial and plot the result. 

```{r}
delay <- flights_tbl %>% 
  group_by(tailnum) %>%
  summarise(count = n(), dist = mean(distance), delay = mean(arr_delay)) %>%
  filter(count > 20, dist < 2000, !is.na(delay)) %>%
  collect

# plot delays
library(ggplot2)
ggplot(delay, aes(dist, delay)) +
  geom_point(aes(size = count), alpha = 1/2) +
  geom_smooth() +
  scale_size_area(max_size = 2)
```

You can write SQL using `DBI` package.

```{r}
library(DBI)

con <- dbConnect(
  RPresto::Presto(),
  host="https://api-presto.treasuredata.com",
  port=443,
  user=Sys.getenv("TD_API_KEY"),
  schema='test',
  catalog='td-presto'
)

flights_preview <- dbGetQuery(con, 'SELECT year, month, day, dep_time, dep_delay, carrier, flight from flights limit 10')
flights_preview
```

You can execute basic TD operation like delete table.

```{r}
delete_table(client, "test", "flights")
```

## References
- [RPresto](https://github.com/prestodb/RPresto)
- [RTD](https://github.com/chezou/RTD)
